\begin{abstract}

Sofware-defined networking (SDN) decouples the control plane from the
forwarding plane in switches, making it possible to create controllers in
software.
%
This has made it easier to build arbitrarily advanced networks, such as
networks that self-optimize for low power-consumption
\cite{Heller:2010:ESE:1855711.1855728}, cloud networks that route live
traffic to moving virtual machines \cite{erickson2008demonstration} as well
as making it possible to research and test new networking protocols using
simulators on a laptop.

A trend in the computing industry is that more services are moved to the
Cloud, reaching a larger number of users.
%
Thus, it becomes more important that these services remain available despite
failure of individual machines, calling for distributed service replication.
%
A huge body of research in the distributed computing community have been
devoted to coming up with protocols that guarantee strong consistency among
the replicas of such a service.
%
The most cited paper on network resilience is the Paxos algorithm
\cite{Lam01} for message ordering.

By extending the OpenFlow protocol, we have implemented steady-phase Paxos
on a software switch, showing how it is possible to compose network
flows that leverage the message ordering guarantees of Paxos as a
constituent element.
%
As a demonstration of its use, we have built a system where distributed
UDP-based services were replicated in-order using Paxos transparently.

We also implemented these Paxos primitives on the switch itself by modifying
Open vSwitch, but were unable to fully test this solution.
%
However, we have demonstrated that by moving primitive operations down to the
switch, it became vastly easier to build advanced flows---compared to
programming them in the controller.

Due to the elegant way in which flows can be built with new OpenFlow
primitives, we believe that our work shows a possible future direction for
the OpenFlow specification that encompass programming the switch.

\end{abstract}
