\section{Capabilities in OpenFlow}
\label{chapter:details.openflow}

Now that we have discussed the main algorithms and our simplification of
them, we must take a look at what OpenFlow can offer us to reach our goals.

\todo{Show ONLY openflow 1.3, possibly 1.4}

To see how we can enable Paxos functionality in OpenFlow, we need to take a
look at what features it can provide us.  There are several versions of the
OpenFlow specification, so we'll review the differences in each
one.\footnote{Unfortunately, the most widely supported version of OpenFlow in
simulators and controllers seem to be OpenFlow version 1.0.}

Naturally, we could implement the whole Paxos algorithm in the controller
itself.  Doing so should be quite trivial: One could modify an existing
implementation and make it use OpenFlow to transmit packets between
switches.  However, that would be very inefficient compared to running the
entire algorithm, or parts of it, on the switch.

Prior to OpenFlow 1.0 \cite{openflow-1.0} there were several working
drafts not meant for implementation.  We will not look at these prior
versions.

In the tables below, you can see what version 1.0 offers in terms of core
functionality.  Some details have been omitted in favor of giving a clear
overview.  For details, see the full specification \cite{openflow-1.0}.

What's most important in 1.0, compared to later versions, is that it only
has \textit{one} flow table and only supports IPv4.  Other than that it has
counters per table, per flow, per port and per queue.  The headers that can
be used for matching packets are listed in table
\ref{table:openflow-1.0.headers} \vpageref{table:openflow-1.0.headers} and
the actions in table \ref{table:openflow-1.0.actions}
\vpageref{table:openflow-1.0.actions}.

By \textit{transport} address and port, we mean \ac{TCP} or \ac{UDP},
depending on what packet is currently matched\index{OpenFlow!transport}.

The specification requires compliant switches to update packet checksums
when modifying fields that require it.

\begin{table}
  \centering
  \begin{tabular}{l}
    \hline
     \textbf{Header-field} \\
    \hline
     Ingress port\index{OpenFlow!match on ingress port} \\

     Ethernet source address\index{OpenFlow!match on Ethernet} \\
     Ethernet destination address\index{OpenFlow!match on Ethernet} \\

     VLAN ID\index{OpenFlow!match on VLAN} \\
     VLAN priority\index{OpenFlow!match on VLAN} \\

     IP source address\index{OpenFlow!match on IP address} \\
     IP destination address\index{OpenFlow!match on IP address} \\
     IP protocol\index{OpenFlow!match on IP protocol} \\
     IP \ac{ToS} bits\index{OpenFlow!match on ToS} \\

     Transport source port\index{OpenFlow!UDP}\index{OpenFlow!TCP}\index{OpenFlow!transport} \\
     Transport destination port \\
    \hline
  \end{tabular}
  \caption{Header-fields that can be matched in OpenFlow 1.0.}
  \label{table:openflow-1.0.headers}
\end{table}
\index{OpenFlow!matching}
\index{OpenFlow!header-fields}
\index{OpenFlow!matching header-fields}

\begin{table}
  \centering
  \begin{tabular}{lll}
    \hline
      \textbf{Action} &
      \textbf{Required} &
      \textbf{Options} \\

    \hline
      Forward\index{OpenFlow!forwarding action} &
      Required &
               To all \\
     & & To controller \\
     & & To local switch \\
     & & To flow table \\
     & & To port \\
    \\
      Forward\index{OpenFlow!forwarding action} &
      Optional &
               Normal \\
     & & Flood\index{OpenFlow!flooding action} \\
    \\
      Enqueue\index{OpenFlow!enqueue action} &
      Optional &
      Can be used to implement, e.g., \acs{QoS}\index{OpenFlow!QoS} \\
    \\
      Drop\index{OpenFlow!drop action} &
      Required &
      Drop packet \\
    \\
      Modify-field\index{OpenFlow!modify-field action} &
      Optional &
               Set or replace VLAN ID\index{OpenFlow!modify VLAN} \\
     & & Set or replace VLAN priority\index{OpenFlow!modify VLAN} \\
     & & Strip any VLAN header\index{OpenFlow!modify VLAN} \\
     & & Replace Ethernet source address\index{OpenFlow!modify Ethernet addresses} \\
     & & Replace Ethernet destination address \\
     & & Replace IPv4 source address\index{OpenFlow!modify IPv4 addresses} \\
     & & Replace IPv4 destination address\index{OpenFlow!modify IPv4 addresses} \\
     & & Replace IPv4 \acs{ToS} bits\index{OpenFlow!modify ToS bits} \\
     & & Replace transport source port\index{OpenFlow!modify transport ports} \\
     & & Replace transport destination port\index{OpenFlow!modify transport ports} \\
    \hline
  \end{tabular}
  \caption{Actions in OpenFlow 1.0.}
  \label{table:openflow-1.0.actions}
\end{table}
\index{OpenFlow!actions}

\todo{Skriv litt mer om features i openflow-versjoner nyere enn 1.0, ikke
  lag tabeller, bare list opp store forskjeller. Bruk dette i
    argumentasjonen under.}

\section{Limitations in OpenFlow}

By looking at what OpenFlow versions 1.1--1.3\index{OpenFlow!versions} offer,
one can see that we can't really make use of any of the added functionality
for running Paxos.  What we need is the ability to run programs on the
switch, which is something OpenFlow does not support at all.  Neither do
their action primitives add up to anything that could be used for
remembering state (such as the current round number) or executing
if-then-else statements.

One possible solution would be to insert a lot of flow table entries that
each waited for a specific round number.  But that would not be an elegant
or practical solution.

They \textit{do}, however, offer us the ability to implement Paxos entirely
on the controller.  We have actually done this, but one of the stated goals
of this thesis was to move parts of the Paxos code down to the switch
itself.  For this we simply need to be able to run full Turing-equivalent
programming languages\index{programming language}.

For remembering state, we looked at the meta-data\index{OpenFlow!metadata}
that is available in later versions of OpenFlow.  However, metadata only
exists as the packet is processed in the pipeline of flow tables, and is
erased when the packet actions are applied at the end.  To remember state,
we will need to add a table to hold such data in the switch.

Finally, we have to look at which OpenFlow versions our software components
support.

Mininet seems to support whatever version of OpenFlow that Open vSwitch uses,
as this is what it uses as a switch.  Open vSwitch\index{Open vSwitch}
supports OpenFlow versions 1.0---1.3\index{Open vSwitch!OpenFlow support}
almost fully, but support for 1.4 is flaky, and may crash.  So 1.4 is out of the question.

The most obvious component to look at is POX\index{POX}, our controller framework in
Python, which only supports OpenFlow 1.0.\footnote{It does seem to support
some \textit{Nicira extensions}\index{Nicira}\index{OpenFlow!Nicira
extensions}, though.  These are extensions that were originally added to
early OpenFlow versions, but much of it has been implemented in later
versions.  There is also a fork of POX (and other software projects) written
by CPqD that adds support for newer OpenFlow versions, but we haven't looked
at it.}\todo{Vær sikker på at vi ikke kan gjøre dette i Nicira extensions.
Mener jeg har rett, men dobbelsjekk.}

But the major point for our decision is what OpenFlow can offer us.
There simply is no way of executing general code, and there is no way to
remember state.\footnote{We even investigated whether we could use the
counters to count round numbers or store them in IPv6 addresses, using VLAN
for storing data, etc.  All those ideas turned out to be very hairy to
implement, with a real possibility of not working correctly.}

All in all, we have decided to use OpenFlow 1.0 where applicable and extend
it where needed.  Using the flow table, controller and bytecode is a simple
but good, practical decision.

\todo{Ta vekk bytecode, forklar at det er irrelevant hva språk/type vi
  bruker}

\section{Decisions}

We have seen that OpenFlow does not offer the capabilities we need to
implement Paxos on the switch.  We could implement Paxos on the controller,
but that would be nearly equivalent to having adding Paxos to the software
running on the end-hosts.\todo{Dette argumentet må være mye sterkere.
  Vi må vise på en mer klar måte at vi ikke kommer utenom programmering.}

Thus, we have decided to use a combination of OpenFlow flow table rules and
programs running in Open vSwitch as bytecode\index{bytecode} compiled
fragments to handle the details of the Paxos
algorithms.\index{OpenFlow!bytecode}

\todo{Legg til selve designet her, siden tittelen sier det! Vis hvor vi
bruker flow tables (ikke vis detaljer, det kommer senere), vis
nettverksflyt, vis hvor bytecode blir eksekvert, og hva som kjører på
controller.}

\section{Choice of Switch Programming Language}

\todo{Ta vekk}

What we want is to provide \textit{simple} primitives that can be
implemented to run custom code \textit{efficiently} on the
hardware---requiring little memory and few cycles per operation---while
still being useful for other networking protocols.
\todo{back up this statement on the last part}

We think this is a good implementation for fastidious hardware implementers,
but \textit{any} programming language---preferably one that can produce
bytecode---would work just as fine.
\todo{Finn ut hvor jeg snakker om Ngaro og flytt tekst over her.}

